Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2022-06-08 inst/manuscript/manuscript_microxanox.Rmd
Update bibliography and finalizing manuscript.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

@@

# \textcolor{red}{TODO}
- **\textcolor{red}{update links}**
- [x] ~~**\textcolor{green}{update cross references}**~~
- [x] ~~**\textcolor{green}{add abstract}**~~
- [x] ~~**\textcolor{green}{check keywords}**~~
- [x] ~~**\textcolor{green}{add more references}**~~
- [x] ~~**\textcolor{green}{expand on Impact section}**~~
- [x] ~~**\textcolor{green}{expand on Conclusion section}**~~
\pagebreak

@@

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.0                                                                         |  
|   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) |  
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | CC BY 4.0                                                                      |  
|   C5    | Code versioning system used                                  | git                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|   C8    | If available Link to developer documentation/manual          | \textcolor{red}{TO BE ADDED}                                                   |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |

@@

We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017. It was with this goal in mind that we developed the *microxanox* package. The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results. The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](LINK NEEDED).

The second stage was to add functionality that would be necessary to answer our research question. Most importantly, we made it possible to have multiple species of bacteria within each of the three functional groups, for the multiple species to differ in their characteristics, and to vary the number of species and amount of variability among them. We also added functionality that allowed: temporally varying environmental conditions, addition of random noise to state variables, and immigration. In addition to the model itself, the package includes some functions to analyse the results as well as visualize the results to provide a starting point for customized visualizations based on own requirements. The basic and additional functionality is described in the package [User Guide](LINK_NEEDED).

@@

A simulation is run using the `run_simulation()` function. In this function, the ODEs are solved using the function `ode()` in the package *deSolve* package [@Soetaert2010]. The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`. The parameter object returned by `new_runsim_parameter()` contains among other things the `strain_parameter` object, which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed please see the *User Guide* which accompanies the package or is available at [User Guide](@LINK_NEEDED)

@@

For a more detailed walk-through of these two approaches and explanation please see the [User Guide](@LINK_NEEDED).

>>>>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  messages = FALSE
)
```

@@

<!-- # \textcolor{red}{TODO} -->
<!-- - [x] ~~**\textcolor{green}{update links}**~~ -->
<!-- - [x] ~~**\textcolor{green}{update cross references}**~~ -->
<!-- - [x] ~~**\textcolor{green}{add abstract}**~~ -->
<!-- - [x] ~~**\textcolor{green}{check keywords}**~~ -->
<!-- - [x] ~~**\textcolor{green}{add more references}**~~ -->
<!-- - [x] ~~**\textcolor{green}{expand on Impact section}**~~ -->
<!-- - [x] ~~**\textcolor{green}{expand on Conclusion section}**~~ -->
\pagebreak

@@

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.0                                                                         |  
|   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
|         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | CC BY 4.0                                                                      |  
|   C5    | Code versioning system used                                  | git                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|   C8    | If available Link to developer documentation/manual          | [Usert Manual](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |

@@

We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017. It was with this goal in mind that we developed the *microxanox* package [@Krug2022]. The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results. The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html).

The second stage was to add functionality that would be necessary to answer our research question. Most importantly, we made it possible to have multiple species of bacteria within each of the three functional groups, for the multiple species to differ in their characteristics, and to vary the number of species and amount of variability among them. We also added functionality that allowed: temporally varying environmental conditions, addition of random noise to state variables, and immigration. In addition to the model itself, the package includes some functions to analyse the results as well as visualize the results to provide a starting point for customized visualizations based on own requirements. The basic and additional functionality is described in the package [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).

@@

A simulation is run using the `run_simulation()` function. In this function, the ODEs are solved using the function `ode()` in the package *deSolve* package [@Soetaert2010]. The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`. The parameter object returned by `new_runsim_parameter()` contains among other things the `strain_parameter` object, which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed please see the *User Guide* which accompanies the package or is available at [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).

@@

For a more detailed walk-through of these two approaches and explanation please see the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).

### 2022-06-08 inst/manuscript/manuscript_microxanox.tex
Update bibliography and finalizing manuscript.

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\usepackage[normalem]{ulem}
% avoid problems with \sout in headers with hyperref:
\pdfstringdefDisableCommands{\renewcommand{\sout}{}}

@@

\hypertarget{section}{%
\section{\texorpdfstring{\textcolor{red}{TODO}}{}}\label{section}}

\begin{itemize}
\tightlist
\item
  \textbf{\textcolor{red}{update links}}
\item[$\boxtimes$]
  \sout{\textbf{\textcolor{green}{update cross references}}}
\item[$\boxtimes$]
  \sout{\textbf{\textcolor{green}{add abstract}}}
\item[$\boxtimes$]
  \sout{\textbf{\textcolor{green}{check keywords}}}
\item[$\boxtimes$]
  \sout{\textbf{\textcolor{green}{add more references}}}
\item[$\boxtimes$]
  \sout{\textbf{\textcolor{green}{expand on Impact section}}}
\item[$\boxtimes$]
  \sout{\textbf{\textcolor{green}{expand on Conclusion section}}}
  \pagebreak
\end{itemize}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule()
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & CC BY 4.0 \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\textcolor{red}{TO BE ADDED} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Rainer@krugs.de \\
\bottomrule()
\end{longtable}

@@

We decided to fill this research gap by making a simulation study of how
within functional group biodiversity affects ecosystem responses to
environmental change \citet{Limberger2022}, and to base our work on the
work and model of \citet{Bush2017}. It was with this goal in mind that
we developed the \emph{microxanox} package. The first stage of
development was to write code from scratch (as there was no available
code to start from) and to confirm that this new implementation would
reproduce the previously published results. The resulting reproduction
is available as one of the package vignettes:
\href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et al}.

The second stage was to add functionality that would be necessary to
answer our research question. Most importantly, we made it possible to
have multiple species of bacteria within each of the three functional
groups, for the multiple species to differ in their characteristics, and
to vary the number of species and amount of variability among them. We
also added functionality that allowed: temporally varying environmental
conditions, addition of random noise to state variables, and
immigration. In addition to the model itself, the package includes some
functions to analyse the results as well as visualize the results to
provide a starting point for customized visualizations based on own
requirements. The basic and additional functionality is described in the
package \href{LINK_NEEDED}{User Guide}.

@@

A simulation is run using the \texttt{run\_simulation()} function. In
this function, the ODEs are solved using the function \texttt{ode()} in
the package \emph{deSolve} package \citep{Soetaert2010}. The
\texttt{run\_simulation()} function needs only one argument - an object
as created by the function \texttt{new\_runsim\_parameter()}. The
parameter object returned by \texttt{new\_runsim\_parameter()} contains
among other things the \texttt{strain\_parameter} object, which can be
created by the function \texttt{new\_strain\_parameter()}. For a
detailed description of the parameter objects, their meaning and how
they are created and have values set and changed please see the
\emph{User Guide} which accompanies the package or is available at
\href{@LINK_NEEDED}{User Guide}

@@

For a more detailed walk-through of these two approaches and explanation
please see the \href{@LINK_NEEDED}{User Guide}.

@@

\begin{verbatim}
#> 
#> Attaching package: 'dplyr'
#> The following objects are masked from 'package:stats':
#> 
#>     filter, lag
#> The following objects are masked from 'package:base':
#> 
#>     intersect, setdiff, setequal, union
#> Warning in knitr::include_graphics(fig): It is highly recommended
#> to use relative paths for images. You had absolute paths: "/
#> Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/
#> gen_uc1_partrep_temporal_state_switching.pdf"
\end{verbatim}

@@

\begin{verbatim}
#> Warning in knitr::include_graphics(fig): It is highly recommended
#> to use relative paths for images. You had absolute paths: "/
#> Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/
#> gen_uc2_user_guide_hysteresis.pdf"
\end{verbatim}

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

\pagebreak

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule()
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & CC BY 4.0 \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{Usert
Manual} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Rainer@krugs.de \\
\bottomrule()
\end{longtable}

@@

We decided to fill this research gap by making a simulation study of how
within functional group biodiversity affects ecosystem responses to
environmental change \citet{Limberger2022}, and to base our work on the
work and model of \citet{Bush2017}. It was with this goal in mind that
we developed the \emph{microxanox} package \citep{Krug2022}. The first
stage of development was to write code from scratch (as there was no
available code to start from) and to confirm that this new
implementation would reproduce the previously published results. The
resulting reproduction is available as one of the package vignettes:
\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
Partial reproduction of Bush et al}.

The second stage was to add functionality that would be necessary to
answer our research question. Most importantly, we made it possible to
have multiple species of bacteria within each of the three functional
groups, for the multiple species to differ in their characteristics, and
to vary the number of species and amount of variability among them. We
also added functionality that allowed: temporally varying environmental
conditions, addition of random noise to state variables, and
immigration. In addition to the model itself, the package includes some
functions to analyse the results as well as visualize the results to
provide a starting point for customized visualizations based on own
requirements. The basic and additional functionality is described in the
package
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

@@

A simulation is run using the \texttt{run\_simulation()} function. In
this function, the ODEs are solved using the function \texttt{ode()} in
the package \emph{deSolve} package \citep{Soetaert2010}. The
\texttt{run\_simulation()} function needs only one argument - an object
as created by the function \texttt{new\_runsim\_parameter()}. The
parameter object returned by \texttt{new\_runsim\_parameter()} contains
among other things the \texttt{strain\_parameter} object, which can be
created by the function \texttt{new\_strain\_parameter()}. For a
detailed description of the parameter objects, their meaning and how
they are created and have values set and changed please see the
\emph{User Guide} which accompanies the package or is available at
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

@@

For a more detailed walk-through of these two approaches and explanation
please see the
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

@@

\begin{verbatim}
#> 
#> Attaching package: 'dplyr'
#> The following objects are masked from 'package:stats':
#> 
#>     filter, lag
#> The following objects are masked from 'package:base':
#> 
#>     intersect, setdiff, setequal, union
\end{verbatim}

@@
